---
title: "R Notebook"
output: html_notebook
---

## Libraries
```{r}

library(tidyverse)
library(lubridate)
library(jsonlite)

library(rvest)
library(httr)
library(eurostat)

library(gghighlight)
library(ggalluvial)

```

## Get & Wrangle Data

GIE-data: https://agsi.gie.eu/#/api
API-key: 21318434befcb49923bb06332c0c571a

(Ik heb een API-key aangemaakt maar ik gebruik die eigenlijk nergens dus ik weet niet of je die nodig hebt of niet.)


```{r}

source("loadGasData.R")

```

## Subsets
```{r}

## last available data for each country
df_gas_latest <- df_gas_storage %>%
  filter(gasDayStartedOn == max(gasDayStartedOn, na.rm = T) & full > 0)

## data for Europe
df_gas_storage_eu <- df_gas_storage %>% 
  filter(country == "EU")

## withdrawal data for Europe
df_gas_withdrawal_eu <- df_gas_storage %>%
  filter(country == "EU" & !is.na(withdrawal)) %>%
  filter(plot_date <= Sys.Date() - 2 & year >= 2015)

```

## GAS STORAGE LEVELS

### EU storage levels

#### Baseplots
```{r}

basePlot <- ggplot() +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#FFDF00"),
    legend.position = "none",
    panel.grid = element_line(colour = "#a7a5a6", size = 0.1),
    panel.grid.minor.y = element_blank(),
    panel.ontop = TRUE
    )

baseLinePlot <- basePlot +
  labs(x = NULL, y = NULL, subtitle = "Yearly gas in storage (%) in the EU") +
  scale_colour_manual(values = c("#a7a5a6","black")) +
  scale_size_manual(values = c(0.2, 1)) +
  scale_x_date(date_labels = "%b", expand = c(0,0)) +
  scale_y_continuous(breaks = seq(0, 100, 25), limits = c(0,100), position = "right", expand = c(0,0))

```

#### Plot EU linegraph
```{r}

## plot eu gas storage graph
baseLinePlot +
  geom_area(
    data = subset(df_gas_storage_eu, year == 2020),
    aes(x = plot_date, y = full_avg * 100), 
    fill = "white") +
  geom_line(
    data = df_gas_storage_eu,
    aes(x = plot_date, y = full * 100, group = year, colour = year >= 2022, size = year >= 2022)) +
  geom_text(
    data = subset(df_gas_storage_eu, gasDayStartedOn == max(gasDayStartedOn, na.rm = T)),
    aes(x = plot_date, y = full * 100, label = format(gasDayStartedOn, "%d/%m/%y")), 
    size = 3, colour = "black", hjust = 0, nudge_y = -5) +
  geom_text(
    data = subset(df_gas_storage_eu, gasDayStartedOn == max(gasDayStartedOn, na.rm = T)),
    aes(x = plot_date, y = full * 100, label = paste0(round(full * 100), "%")), 
    size = 6, colour = "black", hjust = 0, nudge_y = -10, fontface = "bold")

```

#### Storage withdrawals
```{r}

df_gas_withdrawal_eu %>%
  ggplot(aes(x = plot_date, y = cumWithdrawal)) +
  geom_ribbon(aes(ymin = withdrawal_min, ymax = withdrawal_max), alpha = 0.2, fill = "grey") +
  geom_line(aes(y = withdrawal_avg), colour = "grey") +
  geom_line(aes(colour = factor(year))) +
  gghighlight(year >= 2021) +
  scale_x_date(date_labels = "%d %b", expand = c(0,0))

```

#### Projection
When will levels reach the 2015-2020 range?
```{r}

## calculate avg withdrawals based on the 2015-20 range
avg_withdrawals <- df_gas_storage %>%
  group_by(plot_date, country) %>% 
  summarise(
    avg_withdrawal = (mean(withdrawal[year>=2022-5], na.rm = T) - mean(injection[year>=2022-5], na.rm = T))/ 1000,
    withdrawal_min = min(gasInStorage[year>=2015 & year<=2020]),
    withdrawal_max = max(gasInStorage[year>=2015 & year<=2020])
    ) %>% 
  ungroup()

## calculate projection till end of year
df_gas_projection <- df_gas_storage %>% 
  group_by(country) %>% 
  tidyr::complete(gasDayStartedOn = seq.Date(from = max(gasDayStartedOn, na.rm = T), to = ymd("2022-12-31"), by="day")) %>%
  arrange(gasDayStartedOn) %>% 
  fill(year, gasInStorage) %>% 
  mutate(plot_date = as.Date(paste0("2021-", format(gasDayStartedOn,"%m-%d")))) %>% 
  ungroup() %>%
  filter(country %in% c("EU")) %>% 
  filter(year >= 2022) %>%
  select(-c(withdrawal_min, withdrawal_max)) %>%
  left_join(avg_withdrawals, by = c("plot_date","country")) %>% 
  select(country, plot_date, gasDayStartedOn, gasInStorage, avg_withdrawal, withdrawal_min, withdrawal_max) %>% 
  group_by(country) %>% 
  mutate(
    avg_withdrawal = ifelse(gasDayStartedOn > max(df_gas_storage$gasDayStartedOn, na.rm = T), avg_withdrawal, 0),
    category = ifelse(gasDayStartedOn > max(df_gas_storage$gasDayStartedOn, na.rm = T), "projection", "gasInStorage"),
    gasInStorage = ifelse(gasDayStartedOn > max(df_gas_storage$gasDayStartedOn, na.rm = T), gasInStorage - cumsum(avg_withdrawal), gasInStorage),
    diff = gasInStorage - withdrawal_min
      )

## calculate when projection will hit the 2015-20 range
intersection_point <- df_gas_projection %>% 
  slice(0 : which.max(diff >= 0)) %>% 
  filter(gasDayStartedOn == max(gasDayStartedOn))

## plot graph
df_gas_projection %>% 
  ggplot(aes(x = gasDayStartedOn)) +
  geom_ribbon(aes(ymin = withdrawal_min, ymax = withdrawal_max), alpha = 0.5) +
  geom_line(aes(y = gasInStorage, colour = category)) +
  geom_point(data = intersection_point, aes(y = gasInStorage)) +
  geom_text(data = intersection_point, aes(y = gasInStorage, label = gasDayStartedOn), hjust = 0, nudge_x = 2.5) +
  facet_wrap(~country, scales = "free_y", ncol = 1) +
  scale_x_date(date_labels = '%b', breaks = seq.Date(from = ymd("2022-01-01"), to = ymd("2022-12-01"), by = "1 month")) +
  theme(legend.position = "top")

```



### Country storage levels

#### Where are levels below avg...
```{r}

## subset for countries below avg
countries_below <- df_gas_latest %>% 
  filter(diff <= -15)

## filter countries we want to plot
df_plot <- df_gas_storage %>%
  filter(country %in% countries_below$country)

## plot graph
baseLinePlot +
  geom_area(
    data = subset(df_plot, year == 2020),
    aes(x = plot_date, y = full_avg * 100), 
    fill = "white") +
  geom_line(
    data = df_plot,
    aes(x = plot_date, y = full * 100, group = year, colour = year >= 2022, size = year >= 2022)) +
  geom_text(
    data = subset(df_plot, gasDayStartedOn == max(gasDayStartedOn, na.rm = T)),
    aes(x = plot_date, y = full * 100, label = paste0(round(full * 100), "%")), 
    size = 4, colour = "black", hjust = 0, nudge_y = -10, fontface = "bold") +
  facet_wrap(~country_name_nl, scales = "free_x")

```

#### ...same level
```{r}

countries_middle <- df_gas_latest %>% 
  filter(diff > -15  & diff < 15)

## plot
df_plot <- df_gas_storage %>%
  filter(country %in% countries_middle$country)

## plot graph
baseLinePlot +
  geom_area(
    data = subset(df_plot, year == 2020),
    aes(x = plot_date, y = full_avg * 100), 
    fill = "white") +
  geom_line(
    data = df_plot,
    aes(x = plot_date, y = full * 100, group = year, colour = year >= 2022, size = year >= 2022)) +
  geom_text(
    data = subset(df_plot, gasDayStartedOn == max(gasDayStartedOn, na.rm = T)),
    aes(x = plot_date, y = full * 100, label = paste0(round(full * 100), "%")), 
    size = 4, colour = "black", hjust = 0, nudge_y = -10, fontface = "bold") +
  facet_wrap(~country_name_nl, scales = "free_x")

```


#### ...above avg?

```{r}

countries_middle <- df_gas_latest %>% 
  filter(diff >= 15)

## plot
df_plot <- df_gas_storage %>%
  filter(country %in% countries_middle$country)

## plot graph
baseLinePlot +
  geom_area(
    data = subset(df_plot, year == 2020),
    aes(x = plot_date, y = full_avg * 100), 
    fill = "white") +
  geom_line(
    data = df_plot,
    aes(x = plot_date, y = full * 100, group = year, colour = year >= 2022, size = year >= 2022)) +
  geom_text(
    data = subset(df_plot, gasDayStartedOn == max(gasDayStartedOn, na.rm = T)),
    aes(x = plot_date, y = full * 100, label = paste0(round(full * 100), "%")), 
    size = 4, colour = "black", hjust = 0, nudge_y = -10, fontface = "bold") +
  facet_wrap(~country_name_nl, scales = "free_x")

```
### Export
```{r}

## export data for european gas storage levels
df_gas_storage_eu %>% 
  select(plot_date, year, full, full_avg) %>% 
  spread(year, full) %>%
  filter(!is.na(plot_date)) %>% 
  write.csv("export/eu-storage-levels.csv", row.names = F)

## export gas withdrawals for EU
df_gas_withdrawal_eu %>%
  select(plot_date, year, cumWithdrawal, withdrawal_min, withdrawal_max, withdrawal_avg) %>% 
  spread(year, cumWithdrawal) %>% 
  write_csv("export/eu-storage-withdrawals.csv")

```

## CONSUMPTION

### get data
```{r}

library(eurostat)

# df_eurostat_toc
df_monthly_consumption <- get_eurostat("nrg_cb_gasm", type = "label") %>%
  mutate(
    geo = gsub("Germany \\(until 1990 former territory of the FRG\\)","Germany", geo),
    plot_date = lubridate::ymd(paste0("2021-",format(time, "%m-%d"))),
    year = lubridate::year(time)
    ) %>%
  filter(nrg_bal == "Inland consumption - observed") %>%
  filter(unit == "Terajoule (gross calorific value - GCV)") %>% 
  rename(values_tj = values) %>% 
  select(-unit) %>% 
  mutate(values_gwh = values_tj / 3.6)

glimpse(df_monthly_consumption)

```

### average consumption
```{r}

df_avg_consumption <- df_monthly_consumption %>%
  group_by(geo, plot_date) %>%
  summarise(avg_values = mean(values_gwh[year != 2022], na.rm = TRUE)) %>% 
  ungroup()

df_avg_consumption %>%
  filter(geo %in% c("Belgium","France","Netherlands","Germany","European Union - 27 countries (from 2020)")) %>%
  ggplot(aes(x = plot_date, y = avg_values)) +
  geom_area(alpha = 0.2) +
  geom_line(linetype = "11") +
  scale_x_date(date_labels = "%b") +
  facet_wrap(~geo, scales = "free") +
  labs(
    x = "", y = "",
    subtitle = "Monthly consumption natural gas, in GWh",
    caption = "Source: Eurostat"
  ) +
  theme_minimal() +
  theme(legend.position = "top")

```

```{r}

current_storage_values <- df_gas_storage %>% 
  group_by(geo = country_name, country_name_nl) %>% 
  filter(gasDayStartedOn == max(gasDayStartedOn)) %>% 
  ungroup() %>% 
  select(geo, country_name_nl, gasInStorage, full) %>%
  filter(!is.na(full))

winter_consumption <-  df_avg_consumption %>% 
  filter(plot_date == "2021-01-01" | plot_date >= "2021-11-01") %>% 
  group_by(geo) %>% 
  summarise(winterConsumptionGWh = sum(avg_values)) %>% 
  mutate(dailyWinterConsumptionGWh = winterConsumptionGWh/(92)) %>%  # 92 days in nov, dec and jan
  inner_join(current_storage_values, by = "geo") %>% 
  mutate(
    consumptionDays = gasInStorage * 1000 / dailyWinterConsumptionGWh
  )

```

### Winter consumption
```{r}

winter_consumption %>% 
  ggplot(aes(x = reorder(country_name_nl, consumptionDays), y = consumptionDays)) +
  geom_col() +
  geom_text(aes(label = round(consumptionDays, 0)), hjust = 1.1, size = 3, col = "white") +
  coord_flip() +
  labs(
    x = NULL, y = NULL,
    subtitle = "Hoeveel winterdagen kan elk land \nmet zijn huidige voorraden consumeren?"
  )

```
### Export
```{r}

winter_consumption %>% 
  mutate(
    gasInStorage = gasInStorage * 1000,
    `/` = "/",
    `=` = "="
    ) %>% 
  select(LAND = country_name_nl, `RESERVE (GWh)` = gasInStorage,`/`, `GEM. VERBRUIK WINTERDAG (GWh)` = dailyWinterConsumptionGWh,`=`,`AANTAL DAGEN`=consumptionDays) %>% 
  write.csv("export/table.csv", row.names = F)

```


## GAS PRICE

### Plot gas price
```{r}

gas_price %>%
  mutate(ra7_price = runMean(price, 7)) %>% 
  filter(date >= "2021-01-01") %>% 
  ggplot(aes(x = date, y = price)) +
  geom_line(size = 0.2) +
  geom_line(aes(y = ra7_price), colour = "red") +
  scale_y_continuous(limits = c(0,200), expand = c(0,0))

```
## PIPELINES
```{r}

df_pipelines <- read_csv("https://greeninfo-network.github.io/global-gas-infrastructure-tracker/data/data.csv?v=2.1")
  
```

## GAS IMPORT
```{r}
View(eurostat::get_eurostat_toc())
```


```{r}

gas_import <- get_eurostat("nrg_ti_gasm", type = "label") %>% 
  filter(unit == "Million cubic metres" & values > 0)

View(gas_import %>% 
  filter(!(partner %in% c(unique(gas_import$geo), "Total", "Not specified"))) %>% 
  filter(year(time) == 2021) %>% 
  group_by(geo, siec) %>%
  summarise(Total = sum(values, na.rm = T)))

```

```{r}

get_eurostat("nrg_ti_gasm", type = "label") %>% 
  filter(siec == "Natural gas" & year(time) >= 2021 & values > 0) %>%
  spread(unit, values) %>% 
  ggplot(aes(x = `Million cubic metres`, y = `Terajoule (gross calorific value - GCV)`)) +
  geom_point(alpha = 0.2, size = 0.5) +
  geom_smooth(method = "lm", se = F, size = 0.2, colour = "red") +
  scale_x_log10() +
  scale_y_log10()

```

```{r}

bp_regions <- c("North America","S. & Cent. America","Europe","CIS","Middle East","Africa","Asia Pacific")

bp_trade_flows <- readxl::read_xlsx("data/bp-stats-review-2021-all-data.xlsx", sheet = "Gas - Trade movts - pipeline", range = "A3:X40") %>% 
  fill(Continent, .direction = "up") %>% 
  gather("From","Value", -c(To, Continent)) %>%
  filter(Value > 0)

bp_trade_flows_country <- bp_trade_flows %>% filter(!To %in% c(bp_regions,"Total exports"))
bp_trade_flows_regions <- bp_trade_flows %>% filter(To %in% bp_regions)

```

```{r}

library(ggalluvial)

ggplot(data = bp_trade_flows_regions, aes(axis1 = From, axis2 = To, y = Value)) +
  geom_alluvium(aes(fill = To)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("From", "To"), expand = c(0.15, 0.05)) +
  theme_void() +
  theme(legend.position = "none")

```

```{r}

bp_trade_flows_country %>% 
  # filter(Continent %in% c("Europe","North America")) %>%
  filter(From %in% c("Russian Federation","US")) %>%
  ggplot(aes(axis1 = From, axis2 = To, y = Value)) +
  geom_alluvium(aes(fill = Continent), decreasing = TRUE) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("From", "To"), expand = c(0.15, 0.05)) +
  theme(legend.position = "none")

```
## ENTSOG

### How to map Entsog points
```{r}

## base table from Bruegel
entsog_points <- read_xlsx("data/entsog_points.xlsx") %>%
  mutate(id = entsog_point) %>% 
  separate_rows(entsog_point, sep = "; ") %>% 
  group_by(id) %>% 
  mutate(entsog_capacity = entsog_capacity / n())

## filter Russian points
entsog_points_rus <- entsog_points %>% 
  filter(exporting_country == "Russia") %>% 
  mutate(exporting_route = ifelse(is.na(exporting_route), "other", exporting_route))

entsog_points_rus %>% 
  write_json("../webpage/data/throughputMeta.json")

```

### API call

### hidtorical data

```{r}

source("entsogScraper.R")

```

### current month
```{r}

options(timeout = 500)

"https://transparency.entsog.eu/api/v1/operationalData.xlsx?forceDownload=true&pointDirection=es-tso-0006lng-00012entry&from=2022-02-17&to=2022-03-20&indicator=Physical%20Flow,&periodType=day&timezone=CET&limit=-1&dataset=1&directDownload=true"

## Get all gas import points since sept 2021
df_gas_import <- content(GET(paste0("https://transparency.entsog.eu/api/v1/operationalData.csv?forceDownload=true&delimiter=comma&isTransportData=true&dataset=1&from=2022-01-01&to=",Sys.Date(),"&indicator=Physical%20Flow&periodType=day&timezone=CET&periodize=0&limit=-1&idPointType=1")))

## Get all lng import points
df_lng_import <- content(GET(paste0("https://transparency.entsog.eu/api/v1/operationalData.csv?forceDownload=true&delimiter=comma&isTransportData=true&dataset=1&from=2022-01-01&to=",Sys.Date(),"&indicator=Physical%20Flow&periodType=day&timezone=CET&periodize=0&limit=-1&idPointType=2")))

df_entsog_2022 <- bind_rows(df_gas_import, df_lng_import)

```

### Subsets
```{r}

entsog_rus_phys_flow <- df_entsog %>%
  filter(pointLabel %in% entsog_points_rus$entsog_point & directionKey == "entry") %>% 
  left_join(entsog_points_rus, by = c("pointLabel"="entsog_point")) %>%
  mutate(
    date = as.Date(periodFrom),
    value = value / 1000000 # convert kWh to GWh
    ) %>%
  group_by(date, exporting_route) %>% 
  summarise(value = sum(value, na.rm = T))

entsog_phys_by_country <- df_entsog %>% 
  filter(pointLabel %in% entsog_points$entsog_point & directionKey == "entry") %>% 
  left_join(entsog_points, by = c("pointLabel"="entsog_point")) %>%
  mutate(
    date = floor_date(periodFrom, "1 day"),
    value = value / 10000000 # convert kWh to GWh
    ) %>%
  group_by(date, exporting_country) %>% 
  summarise(value = sum(value, na.rm = T))

entsog_rus_phys_flow_weekly <- entsog_rus_phys_flow %>% 
  mutate(date_floor = floor_date(date, "1 week")) %>% 
  group_by(exporting_route, date_floor) %>% 
  summarise(value =  sum(value))

```

### Export
```{r}

entsog_rus_phys_flow_weekly %>%
  filter(date_floor >= "2021-10-01" & date_floor < floor_date(Sys.Date(), "1 week") & value > 0) %>%
  rename(category = exporting_route) %>% 
  jsonlite::write_json("../webpage/src/data/bars.json")

entsog_rus_phys_flow %>%
  filter(date < Sys.Date() & date >= "2021-12-01") %>% 
  mutate(value = value * 100) %>%
  spread(exporting_route, value) %>% 
  write_csv("dw-russian-export.csv")

```

### Graphs

### Russian gas (weekly)
```{r}

entsog_rus_phys_flow_weekly %>% 
  filter(date_floor >= "2021-10-01" & date_floor < floor_date(Sys.Date(), "1 week") & value > 0) %>%
  ggplot(aes(x = date_floor, y = value)) +
  geom_col() +
  facet_wrap(~exporting_route, scales = "free")

```

#### Russian gas
```{r}

entsog_rus_phys_flow %>%
  # filter(exporting_route %in% c("yamal","nordstream","ukraine")) %>%
  filter(date <= Sys.Date() - 2 & date >= "2021-10-01") %>%
  ggplot(aes(x = date, y = value)) +
  geom_area(aes(fill = exporting_route)) +
  scale_x_date(breaks = seq.Date(ymd(floor_date(min(entsog_rus_phys_flow$date), "1 month")), ymd(floor_date(max(entsog_rus_phys_flow$date), "1 month")), "2 month"), date_labels = "%b/%y")

```
#### Eu gas import
```{r}

bump_data <- entsog_phys_by_country %>% 
  filter(date < Sys.Date() - 1) %>%
  mutate(
    date = floor_date(date, "1 week"),
    exporting_country = ifelse(exporting_country %in% c("Azerbaijan","Libya"), "Other",exporting_country)
  ) %>% 
  group_by(exporting_country, date) %>%
  summarise(value = sum(value)) %>%
  ungroup() %>%
  filter(date < floor_date(Sys.Date(), "1 week") & date >= "2021-10-01")

bump_data %>%
  ggplot(aes(x = date, y = value, alluvium = exporting_country)) +
  geom_alluvium(aes(fill = exporting_country, colour = exporting_country),
                alpha = .75, decreasing = FALSE) +
  scale_colour_brewer(type = "qual", palette = "Paired") +
  scale_fill_brewer(type = "qual", palette = "Paired") +
  theme_minimal()

bump_data %>%
  ggplot(aes(x = date, y = value, colour = exporting_country, fill = exporting_country)) +
  geom_area() +
  scale_colour_brewer(type = "qual", palette = "Paired") +
  scale_fill_brewer(type = "qual", palette = "Paired") +
  theme_minimal()

```

#### LNG gas import
```{r}

df_entsog %>%
  filter(pointLabel %in% entsog_points$entsog_point & directionKey == "entry") %>% 
  left_join(entsog_points, by = c("pointLabel"="entsog_point")) %>%
  mutate(
    date = floor_date(periodFrom, "1 day"),
    value = value / 1000000
    ) %>% 
  filter(exporting_country == "LNG" & !is.na(value)) %>% 
  arrange(date) %>% 
  group_by(pointLabel, date) %>%
  summarise(value = sum(value)) %>% 
  ungroup() %>%
  group_by(pointLabel) %>% 
  mutate(n = n()) %>% 
  filter(n>7) %>%
  filter(date >= "2022-01-01") %>% 
  mutate(ra7_value = TTR::runMean(value, 7)) %>%
  ggplot(aes(x = date, y = value, colour = pointLabel, fill = pointLabel)) +
  # geom_point(aes(y = value), size = 0.5, alpha = 0.3) +
  geom_line() +
  gghighlight(pointLabel %in% c("Gate Terminal (I)"), use_direct_label = F)

```
```{r}

df_plot <- df_entsog %>%
  filter(pointLabel %in% entsog_points_rus$entsog_point & directionKey == "entry") %>% 
  left_join(entsog_points_rus, by = c("pointLabel"="entsog_point")) %>%
  mutate(
    date = as.Date(periodFrom),
    value = value / 10000000
    ) %>% 
  group_by(date, exporting_route) %>% 
  summarise(
    value = sum(value, na.rm = T),
    entsog_capacity = sum(entsog_capacity)/10)

df_plot %>%
  ggplot(aes(x = date, y = value, fill = exporting_route)) +
  geom_col(width = 1) +
  # geom_line(aes(y = entsog_capacity)) +
  facet_wrap(~exporting_route, scales = "free")

```

```{r}

entsog_phys_by_country %>% 
  arrange(date) %>% 
  group_by(exporting_country) %>%
  mutate(ra7_value = runMean(value, 14)) %>%
  filter(date >= "2021-01-01" & date < Sys.Date() - 2) %>%
  ggplot(aes(x = date, y = value, colour = exporting_country)) +
  geom_point(size = 0.1) +
  geom_line(aes(y = ra7_value))

entsog_phys_by_country %>% 
  filter(date >= "2021-01-01" & date < Sys.Date() - 2) %>% 
  arrange(date) %>% 
  group_by(exporting_country) %>% 
  mutate(ra7_value = runMean(value, 14)) %>%
  ungroup() %>% 
  gather("category","value", value, ra7_value) %>% 
  mutate(category = paste(exporting_country, category)) %>% 
  select(date, category, value) %>%
  spread(category, value) %>% 
  write_csv("entsog_phys_by_country.csv")

```
```{r}

df_lng_points <- read.csv("data/IGGIN/data/IGGIN_LNGs.csv", sep = ";") %>%
  select(name, lat, long)

lng_points <- jsonlite::fromJSON("data/IGGIN/data/IGGIN_LNGs.geojson")

df_lng_points <- bind_cols(df_lng_points, lng_points$features$properties$param)

df_lng_points %>% 
  write_csv("lng-points.csv")

```

```{r}

lng_flow <- df_entsog %>%
  filter(pointLabel %in% entsog_points$entsog_point & directionKey == "entry") %>% 
  left_join(entsog_points, by = c("pointLabel"="entsog_point")) %>%
  mutate(
    value = value / 1000000,
    pointLabel = toupper(pointLabel)
    ) %>% 
  filter(exporting_country == "LNG" & !is.na(value) & periodFrom >= "2022-01-01") %>%
  select(pointLabel, periodFrom, value) %>% 
  group_by(pointLabel) %>% 
  summarise(avgFlow = mean(value))

df_lng_points %>% 
  mutate(name = toupper(name)) %>% 
  left_join(lng_flow, by = c("name"="pointLabel")) %>% 
  ggplot(aes(x = reorder(name, max_cap_store2pipe_M_m3_per_d), y = max_cap_store2pipe_M_m3_per_d / 10.55 * 1000)) +
  geom_col() +
  geom_col(aes(y = avgFlow * 10), fill = "red", alpha = 0.6) +
  coord_flip()

lng_flow %>% left_join(df_lng_points %>%   mutate(name = toupper(name)), by = c("pointLabel"="name"))

```

```{r}

entsog_phys_by_country %>%
  mutate(date = floor_date(date, "1 week")) %>% 
  filter(year(date) >= 2021 & date < floor_date(Sys.Date() - 2)) %>%
  group_by(date, exporting_country) %>% 
  summarise(value = sum(value, na.rm = T)) %>% 
  ungroup() %>%
  select(date, value, exporting_country) %>% 
  # spread(exporting_country, value) %>%
  # write_csv("dw-export-country.csv")
  ggplot(aes(x = date, y = value, fill = exporting_country)) +
  geom_area(position = "fill")

```



```{r}

df_entsog %>% 
  filter(pointLabel %in% entsog_points$entsog_point & directionKey == "entry") %>% 
  left_join(entsog_points, by = c("pointLabel"="entsog_point")) %>%
  mutate(
    plot_date = as.Date(paste0("2022-", format(as.Date(periodFrom),"%m-%d"))),
    plot_date = floor_date(plot_date, "1 week") + 2,
    year = year(periodFrom),
    value = value / 1000000
    ) %>% 
  filter(periodFrom < floor_date(Sys.Date(), "1 week")) %>% 
  group_by(year, exporting_country, plot_date) %>%
  summarise(value = sum(value, na.rm = T)) %>%
  ungroup() %>% 
  filter(!is.na(exporting_country)) %>%
  ggplot(aes(x = plot_date, y = value, colour = as.factor(year))) +
  geom_line() +
  gghighlight(year >= 2021, calculate_per_facet = T) +
  facet_wrap(~exporting_country, scales = "free_y")

```

```{r}

View(df_lng_import %>% 
  filter(directionKey == "entry") %>%
  mutate(id = gsub("(.{37})|kWh/d", "", id)) %>% 
  distinct(pointLabel, id) %>% 
  arrange(pointLabel))

```


## EU TARGETS
```{r}

last_data <- df_gas_storage %>% 
  group_by(country) %>% 
  filter(gasDayStartedOn == max(gasDayStartedOn)) %>% 
  select(country, date = plot_date, target = full)

gas_capacity <- df_gas_storage %>% 
  group_by(country) %>% 
  filter(gasDayStartedOn == max(gasDayStartedOn)) %>% 
  distinct(country, workingGasVolume)

gas_targets <- read_csv("data/eu-targets.csv") %>% 
  gather("date","target",-c(country)) %>% 
  mutate(
    date = ymd(date),
    target = as.numeric(gsub("%","",target)) / 100,
    ) %>% 
  bind_rows(last_data) %>% 
  left_join(gas_capacity) %>% 
  mutate(targetGasVolume = target * as.numeric(workingGasVolume))

gas_targets

```



```{r}

selected_countries <- c("EU","BE","DE")

df_gas_storage %>% 
  filter(country %in% selected_countries) %>% 
  filter(plot_date >= "2021-03-01") %>%
  ggplot(aes(x = plot_date, y = full)) +
  geom_ribbon(aes(ymin = full_min, ymax = full_max), alpha = 0.1) +
  geom_line(aes(y = full_avg), linetype = "11", colour = "darkgrey") +
  geom_line(aes(colour = factor(year))) +
  scale_colour_manual(values = c("darkgrey","black")) +
  gghighlight(year >= 2021, use_direct_label = F, calculate_per_facet = T) +
  geom_hline(yintercept = 0, size = 0) +
  geom_line(data = subset(gas_targets, country %in% selected_countries),
            aes(x = date, y = target), linetype = "11", colour = "red") +
  geom_point(data = subset(gas_targets, country %in% selected_countries),
            aes(x = date, y = target), colour = "red", size = 0.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  facet_wrap(~country, scales = "free_y", ncol = 2) +
  theme_minimal() +
  theme(legend.position = "top") +
  labs(x = NULL, y = NULL, subtitle = "jaarlijks verloop vulgraad en target voor 2022")

```


